---
title: "simpleRegress"
author: Sarah Castellon
date: October 15 2024
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted r square for the LENGTH~AGE is .07476

## Model LENGTH ~ AGE  this is an example  

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square for the LENGTH~AGE is .05073

## Model LENGTH ~ HEADLEN  this is an example  

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square for the LENGTH~HEADLEN is 0.8434

## Model LENGTH ~ HEADWTH  this is an example    

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted r square for the LENGTH~HEADWTH is 0.5321

## Model LENGTH ~ CHEST this is an example  

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted r square for the LENGTH~CHEST is 0.7869

## Model LENGTH ~ WEIGHTthis is an example  

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted r square for the LENGTH~WEIGHT is 0.7423.

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better.

In comparison of all the models other then NECK from worst to best is as follows...  
LENGTH~AGE r squared is 0.5073  
LENGTH~HEADWTH r squared is 0.5321    
LENGTH~WEIGHT r squared is 0.7423  
LENGTH~CHEST r squared is 0.7869    
LENGTH~HEADLEN r squared is 0.8434    





